---
title: "Learn_R"
author: "Jessie Berta-Thompson"
date: "February 8, 2018"
output: html_document
---
##R
Use this document to run examples from "The Book of R" so that I can do more intentional R learning. 
30 minutes per gardens day.

This lives in a git repository on the DBG Q drive under "Learn_R/" - so push it at the end of each day's progress!

** Do this at beginning of sessions in this workbook **
Copy this to command line (get working directory in the right place - check console window header to make sure)
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Learn_R/")
Clear out workspace
```{r}
rm(list = ls())
```
** End of warmup **



#Section 1.2 Opening R for the first time (page 4)
Customize the prompt?
you can replace the prompt in the console ">" with anything
The Book of R likes to change it to ">R "
I might like project-specific or encouraging ones.
```{r}
options(prompt="R> ")
options(prompt = "Hooray!")
#I think I like vanilla for now, but I like the idea of positivity and customization! could be useful when using multiple code platforms & taking notes.
options(prompt="> ")
```

Up to a previous line of code on command line - arrow up, and arrow down to get back to more recent. 
Ctrl-shift-enter runs the whole chunk of code.
Ctrl-enter runs just line with cursor, or a highlighted section.

** skipped to Rstudio appendix of book pg 784 to get more customization tricks (main text uses base R console/editor)
Tools -> Global Options 

# how to comment in R
```{r}
#comments can go here
1+1 #comments can go here

```

# Working directory
This is where it looks for files on import and saves to, unless path specificied.
```{r}
#Check and see what the working directory is
getwd()

#If you've just opened a .Rmd it seems to set the wd to where it resides.

#Change the working directory to one level higher than where this lives
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/")

getwd()

setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Learn_R")

```

This is slightly complicated in chunks. on the whole, just use paths carefully. 
Sometimes fine to use working directory if all the contents of a project in one place
When more complicated, better to specify paths. 
Good to check in on working directory to make sure you're where you think you are.


Rstudio - there's supposed to be something like this to change working directory for this chunk and all others subsequent to run.
```{r setup}
knitr::opts_knit$set(root.dir = "Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/")

```

but I can't get it to work. 

What does work is entering command in the console directly. After doing that, it applies to chunks. 
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/")
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Learn_R/")
Added this to beginning of document as warmup.

Usefully, the current working directory is shown at the top of the Console pane in Rstudio.
You can also change it in gui menu Session -> Set Working Directory

#Installing and loading R packages (1.2.4 pg 7)
There are lots of built-in commands for calculations, stats and plotting.
There are lots more in packages or libraries.
Some packages come with the base R distribution (no need to install) but need to be loaded for use/access to functions by calling library():
```{r}
library("MASS")
```
Packages need to be reloaded for each R session.

Other packages need to be downloaded and installed, from a repository, usually CRAN (The Comprehensive R Archive Network).
With an internet connection, this can be done with install.packages().
ks (kernel smoother, for multivariate and univariate data) will be used later in this book. 
Executing the install command will connect to local CRAN mirror and download and install the ks package and other pakcages it uses (dependencies)
```{r}
install.packages("ks")

```
This step you only have to do once. Notice quotes in commands above - feed it name of package as string.
Now load the new package
```{r}
library("ks")

```
That one threw a warning/message:
package ???ks??? was built under R version 3.4.3

Which made me curious about my current R version. Check on that with this command:

```{r}
version

```

There are newer versions from:
https://www.r-project.org/
current is:
3.4.3
But I had trouble installing without permissions. 
Try again?
If I ignore one warning about associating Rdata with R version, then it works. 
To get R studio to see it, you have to restart Rstudio. (did that, it worked! we got 3.4.3)

To keep everything in best possible working order, a single command can be used to check all installed packages for more recent updates:
```{r}
update.packages()
```

This took awhile - my first time running it on this computer. 
You have to look at console and respond (y) to updating each package.
(not sure what c in y/n/c response options means)

pick up next time on page 8, 1.2.5 help files. 
2.23.18

R is full of help files (e.g. to understand function arguments)
There are a few ways to get at them. 

```{r}
#these are the same. ? is a short cut for help. This brings up a help file (a different panel in Rstudio, help window in base R console)
#this acts on a specific known function to pull the help file
help(mean)
?mean

#If you don't know the specific function, but want to search help in English across all installed packages
# pass a string with search term of interest to help.search()
help.search("mean")
#this one also has a shortcut
??"mean"

#results are listed as library::function
```

Function help files usually include a 
description of what it does
usage on how to set up arguments and what defaults are -  shown with "=" to set. 
arguments - info about what you pass to the function
value - info on what function returns
references
related help files
examples - often you can copy and paste them into console to see what they do
Warnings - common mistakes
Details - more
Notes - more

These are hard to read at first, but book of R says to try learning to read them easily!
So, look at a few more...
```{r}
#install.packages("ggplot2")
#library(ggplot2)
??"ggplot"
?gc
gc(verbose=getOption("verbose"))
gcinfo(TRUE)
?read.table
```

Then the book does a shoutout to RStudio
it has a "direct-submission code editor"
Might be nice to learn about standalone R code files sometime, in addition to this markdown way of working. 
See Rstudio appendix (could work through that section soon). Also need to figure out knitting, to save these with results in them. Right now I'm using these Rmds to produce other files, but they aren't saving their work (no code output or plots save)

left off at section 1.3 pg 11, Saving work and Exiting R (and picked up here 3-5-18)

## Saving and Exiting R
when you're done working, think about:
1) R objects you've made
2) R script files in editor

You can save all objects (assigned) in current session to a 'workspace image' in a .Rdata
In Rstudio GUI, this lives at:
Session -> Save Workspace AS

Loading a saved workspace is a way to pick up where you left off. 
save.image() saves workspace image (code version)
load() brings the .Rdata into the current workspace

What's in active workspace?
ls() shows all objects, variables and user-defined functions
```{r}
ls()
```
#quit R:
```{r}
q()
```
This brings up a dialog about saving.
It asks if you want to save a workspace image - writes an unnamed .Rdata in current directory with workspace image.
If there's an unnamed .Rdata in default working directory (on main gardens compter, dafault when you open Rstudio is ~/), 
that workspace will be automatically opened when you start. It says what it did/where it got workspace image from in the console welcome.

With the workspace .Rdata, R also saves a file with line-by-line history of commands - supports scrolling through past commands (.Rhistory)
To see these, look at files in working directory, including invisibles, with this:
```{r}
dir(all.files = TRUE)

```

Save editor scripts at .R files (helps operating system connect them with R).
Scripts more useful for saving work than workspace (can always re-run!)

#Conventions 1.4.1 Coding
Long lines can be split at commas in the middle of commands (without actually splitting line) with return, and they will be indented to the level of the opening parentheses of the relevant (split) command/function
```{r}
ordfac.vec <- factor(x = c("Small", "Large", "Large", "Regular", "Small"), 
                     levels = c("Small", "Regular", "Large"), 
                     ordered = TRUE)
```

Left off at 1.4.2 Conventions - Math and Equation References


