---
title: "Learn_R"
author: "Jessie Berta-Thompson"
date: "February 8, 2018"
output: html_document
---
##R
Use this document to run examples from "The Book of R" so that I can do more intentional R learning. 
30 minutes per gardens day.

This lives in a git repository on the DBG Q drive under "Learn_R/" - so push it at the end of each day's progress!

#Section 1.2 Opening R for the first time (page 4)
Customize the prompt?
you can replace the prompt in the console ">" with anything
The Book of R likes to change it to ">R "
I might like project-specific or encouraging ones.
```{r}
options(prompt="R> ")
options(prompt = "Hooray!")
#I think I like vanilla for now, but I like the idea of positivity and customization! could be useful when using multiple code platforms & taking notes.
options(prompt="> ")
```

Up to a previous line of code on command line - arrow up, and arrow down to get back to more recent. 
Ctrl-shift-enter runs the whole chunk of code.
Ctrl-enter runs just line with cursor, or a highlighted section.

** skipped to Rstudio appendix of book pg 784 to get more customization tricks (main text uses base R console/editor)
Tools -> Global Options 

# how to comment in R
```{r}
#comments can go here
1+1 #comments can go here

```

# Working directory
This is where it looks for files on import and saves to, unless path specificied.
```{r}
#Check and see what the working directory is
getwd()

#If you've just opened a .Rmd it seems to set the wd to where it resides.

#Change the working directory to one level higher than where this lives
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/")

getwd()

setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Learn_R")


```

This is slightly complicated in chunks. on the whole, just use paths carefully. 
Sometimes fine to use working directory if all the contents of a project in one place
When more complicated, better to specify paths. 
Good to check in on working directory to make sure you're where you think you are.


Rstudio - there's supposed to be something like this to change working directory for this chunk and all others subsequent to run.
```{r setup}
knitr::opts_knit$set(root.dir = "Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/")

```

but I can't get it to work. 

What does work is entering command in the console directly. 
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/")
setwd("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Modelling/Learn_R/")

Usefully, the current working directory is shown at the top of the Console pane in Rstudio.
You can also change it in gui menu Session -> Set Working Directory